---
title: "assignment_07_MunjewarSheetal"
author: "Sheetal M"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
  word_document: default
#bibliography: bibliography.bib
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Install and Load required packages :
```{r, echo=FALSE}
# Package names
# packages <- c("ggplot2","dplyr","tidyr","magrittr","tidyverse","purrr")
packages <- c("ggplot2","dplyr","magrittr","tidyverse","purrr","pander","pandoc")

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))
```

## Set the working directory to the root of your DSC 520 directory
setwd("E:\\Data_Science_DSC510\\DSC520-Statistics\\dsc520") 

## Load the `data/r4ds/heights.csv` to
heights_df <- read.csv("data/r4ds/heights.csv") 


## Using `cor()` compute correlation coefficients for
```{r}
## Set the working directory to the root of your DSC 520 directory
setwd("E:\\Data_Science_DSC510\\DSC520-Statistics\\dsc520") 

## Load the `data/r4ds/heights.csv` to
heights_df <- read.csv("data/r4ds/heights.csv")

str(heights_df)
### height vs. earn
cor(heights_df$earn,heights_df$height, method='pearson')
### age vs. earn
cor(heights_df$age,heights_df$earn, method='spearman' )

### ed vs. earn
plot(heights_df$earn ~ heights_df$ed,
     data = heights_df,
     main = "Education vs Earning",
     xlab = "Education",
     ylab = "Earning")

cor(heights_df$ed,heights_df$earn, method = 'kendall')
cor(heights_df$ed,heights_df$earn, method = 'pearson')

# cor(heights_df)
plot(heights_df, main = "Height vs Earning")
```


### Spurious correlation
### The following is data on US spending on science, space, and technology in millions of today's dollars
### and Suicides by hanging strangulation and suffocation for the years 1999 to 2009
### Compute the correlation between these variables
```{r}

tech_spending <- c(18079, 18594, 19753, 20734, 20831, 23029, 23597, 23584, 25525, 27731, 29449)
suicides <- c(5427, 5688, 6198, 6462, 6635, 7336, 7248, 7491, 8161, 8578, 9000)
ss_df <- data.frame(tech_spending,suicides)
str(ss_df)
cor(ss_df)
plot(ss_df, main = "spending vs sucide")
cor(ss_df)
cov(ss_df)

#calculate Pearson correlation coefficient and ignore any rows with NA
cor(ss_df$tech_spending,ss_df$suicides, use='complete.obs')

```
